Variant{\sc Spark}: Population Scale Clustering of Genotype Information

### 2015-08-14 doc/publications/journal2014/paper.tex

\subsection*{Variant{\sc Spark} allows genome wide sampling of variants to improve clustering quality}
Add the results for the scalling from 1\% to X\%

>>>>

\subsection*{Variant{\sc Spark} allows genome wide sampling of variants to improve clustering quality}

To demonstrate the scalability of Spark when applied to clustering individuals on their variants, we scale the size up from chromosome 22 to all 22 autosomes. Although this pre-processing and clustering is successful, the memory required for clustering increases dramatically.

Where our previous clustering of chromosome 22 required no more than 4GB memory per executor, clustering the entire genome increases this requirement to over 20GB. On the other hand, the memory requirements for our pre-processing implementation remain unchanged, at 2 gigabytes.
One factor which results in these differing memory requirements is the sparsity of the vectors used to store the variants. Due to their sparse nature (the abundance of 0s) we can store them efficiently as Sparse Vectors, rather than Dense Vectors. However, to cluster these vectors, the k-means algorithm converts them to Dense Vectors.
With Dense Vectors, all the 0 values (which are excluded from Sparse Vectors so don't require any memory) are explicitly defined, considerably increasing the memory required.
